---
title: "Prepping for dada2"
author: "Vicki Hertzberg and Madelyn Houser"
date: "2/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To work along with the dada2 lecture on 2/16, you will need to install a number of programs. Most of these come from Bioconductor.

You will need the following programs from CRAN:
  
ggplot2
  
```{r ggplot2}
#Remove comment symbol from next line for it to work. I am keeping the next line commented, however, because I do not need to install ggplot2; it is already installed.

#install.packages("ggplot2")
```

  
You will need to install the following programs from Bioconductor:
  
dada2  
phyloseq  
ShortRead  
DECIPHER  
  
To install a program from Bioconductor, you will first need to install Bioconductor. To do this, execute the following code IN THE CONSOLE WINDOW:
  
  
```{r BiocManager}
#Remove comment symbol from next lines for it to work. I am keeping the next lines commented, however, because I do not need to install BiocManager; it is already installed.

#if (!requireNamespace("BiocManager", quietly = TRUE))  
#    install.packages("BiocManager")  
#BiocManager::install(version = "3.14")  
```
  
  
Next install the other programs. For instance to install dada2, use the following code:

```{r dada2}
#Remove comment symbol from next line for it to work. I am keeping the next line commented, however, because I do not need to install dada2; it is already installed.

#BiocManager::install("dada2")

```

To install the other programs, use a similar chunk of code, this time substituting the package name for dada2.

You can check on the package version of all of the programs you have installed.

```{r versions, results='hide'}

packageVersion("ggplot2")
packageVersion("dada2")
packageVersion("phyloseq")
packageVersion("DECIPHER")
packageVersion("ShortRead")

```

There are some other files you will need to download.

First you wil need to download the mothur MiSeq SOP dataset that you can find at http://mothur.s3.us-east-2.amazonaws.com/wiki/miseqsopdata.zip

Unzip this download and store the folder on your hard drive (not your desktop), and remember the path to it. You should have 45 files. 

Also, go to https://zenodo.org/record/3986799#.YC7mEi2cbUI and download:  
silva_nr99_v138_train_set.fa.gz  
silva_nr99_v138_wSpecies_train_set.fa.gz  
silva_species_assignment_v138.fa.gz  

Store these files in the same place where you are storing the mothur example dataset. DO NOT UNZIP.

We are now ready to take on dada2 on 2/16/22.



